Negative Selection in the 20th Century

### 2020-05-19 why-negative-selection.Rmd
regressions on age of parents at own birth

* Divide siblings regressions by own YOB, into early and late.
  - Use this to confirm the basic "selection decreasing" story. DONE.
  - Actually almost no change.
  - Overall, diffs between siblings and children regressions
    now look more like ascertainment bias than change over time.
  - (I checked my code: I hadn't been controlling for one main term when
    running the interaction. Doh.)

* Work more on weighting the data?
  - weight by age at first live birth
  - education using ISCED

* direct regressions of scores on age_flb

@@

@@

loadd(res_together)
n_regs <- as.double(nrow(res_together))
res_together %>% 
      mutate(
        term = pretty_names(term),
        term = fct_reorder(term, estimate, order_abs(2)),
        dep.var = dplyr::recode(dep.var, 
          "n_children" = "N Children", "n_sibs" = "N Siblings"),
        dep.var = fct_relevel(dep.var, "N Siblings")
      ) %>% 
      ggplot(aes(y = term, x = estimate, color = p.value < 0.05/{{n_regs}}, 
        fill  = PCs)) + 
        pval_geom_point(alpha = 0.8) +
        my_vline +
        facet_wrap(vars(dep.var)) +
        pval_scale() +
        my_fill_scale(n = 2, direction = -1) +
        theme(legend.position = "bottom")
```

@@

n_regs <- as.double(nrow(res_sex))
res_sex %>% 
      mutate(
        term = pretty_names(term),
        term = fct_reorder(term, estimate, order_abs(1))
      ) %>% 
      ggplot(aes(y = term, x = estimate, fill = sex, 
            colour = p.value < 0.05/{{n_regs}})) +
        my_vline +
        pval_geom_point() +
        pval_scale() +
        my_fill_scale(n = 2, direction = -1)

@@

res_combined %>% 
      mutate(
        score_name = fct_reorder(score_name, estimate, order_abs(2)),
        Regression = ifelse(reg.type == "raw", "Bivariate", "Controlled for age FLB")
      ) %>% 
      ggplot(aes(y = score_name, x = estimate, fill = Regression, 
        color = p.value < 0.05/{{n_regs}})) + 
        geom_vline(xintercept = 0, linetype = "dashed", color = "grey20") +
        pval_geom_point() +
        pval_scale() +
        scale_fill_brewer(type = "qual", palette = 3)
```

@@

The dataset has no information on fathers' age at birth of their first child.
However, we can calculate this information for the parents' generation,
for the subsets of respondents who reported their mother's or father's age
and who had no elder siblings. We run *sibling* regressions on these subsets,
controlling for either parent's age at their birth.
Figure \@ref(fig:plot-age-birth-parents) shows
the results. Effect sizes are very similar, whether controlling for
father's or mother's age at respondent's birth or mother's age at respondent's birth. Unlike for the respondent's own generation, effect sizes are positively correlated with
the effect sizes from bivariate regressions (father's age at birth: 
$\rho$ `r cor_raw_fath`; mother's age at birth: $\rho$ `r cor_raw_moth`).

```{r plot-age-birth-parents}

@@

res_age_birth_parents %>% 
    filter(! term %in% c("fath_age_birth", "moth_age_birth")) %>% 
    mutate(
      term    = pretty_names(term),
      term    = fct_reorder(term, estimate, order_abs(2)),
      Control = ifelse(control == "fath_age_birth", 
                  "Father's age at birth", 
                  "Mother's age at birth"
                )
    ) %>% 
    ggplot(aes(y = term, x = estimate, fill = Control, 
      colour = p.value < 0.05/{{n_regs}})) + 
      pval_geom_point(alpha = 0.8) +  
      my_vline +
      my_fill_scale(2) +
      pval_scale()

@@

n_regs <- as.double(nrow(res_partners))
res_partners %>% 
      mutate(
        score_name = pretty_names(score_name),
        score_name = fct_reorder(score_name, estimate, order_abs(1))
      ) %>% 
      ggplot(aes(y = score_name, x = estimate, fill = `N partners`, 
        colour = p.value < 0.05/{{n_regs}})) + 
        my_vline +
        pval_geom_point() +
        facet_wrap(vars(sex)) +
        my_fill_scale(n = 2) +
        pval_scale() +
        theme(legend.position = "bottom")

@@

drake::loadd(res_edu)
n_regs <- as.double(nrow(res_edu))
res_edu %>% 
      mutate(
        term = pretty_names(term), 
        term = fct_reorder(term, estimate, order_abs(1)),
        "Age left FTE" = fct_relevel(age_fte_cat, "< 16", "16-18", "> 18")
      ) %>% 
      ggplot(aes(y = term, x = estimate, colour = p.value < 0.05/{{n_regs}}, 
        fill = `Age left FTE`)) +
        my_vline +
        pval_geom_point() + 
        pval_scale() +
        my_fill_scale(n = 3, direction = -1)

@@

n_regs <- as.double(nrow(res_income))
res_income %>% 
      mutate(
        term = pretty_names(term), 
        term = fct_reorder(term, estimate, order_abs(1)),
        Income = factor(income_cat, 
          labels = c("< £18K", "£18-30K", "£31-51K", "£52-100K", "> £100K")
        )
      ) %>% 
      ggplot(aes(y = term, x = estimate, fill = Income, 
        colour = p.value < 0.05/{{n_regs}})) +
        my_vline +
        pval_geom_point(alpha = 0.8) + 
        my_fill_scale(n = 5, direction = -1) +
        pval_scale()
```

>>>>

* Work more on weighting the data?
  - weight by age at first live birth
  - education using ISCED?
  - geography?
  - ask for existing work by others?

@@

my_labs <- function (x = "Effect size", y = "", ...) labs(x = x, y = y, ...)

@@

loadd(res_together)
n_regs <- as.double(nrow(res_together))
res_together %>% 
      mutate(
        term = pretty_names(term),
        term = fct_reorder(term, estimate, order_abs(2)),
        dep.var = dplyr::recode(dep.var, 
          "n_children" = "N Children", "n_sibs" = "N Siblings"),
        dep.var = fct_relevel(dep.var, "N Siblings")
      ) %>% 
      ggplot(aes(y = term, x = estimate, color = p.value < 0.05/{{n_regs}}, 
        fill  = PCs)) + 
        pval_geom_point(alpha = 0.8) +
        my_vline +
        facet_wrap(vars(dep.var)) +
        pval_scale() +
        my_fill_scale(n = 2, direction = -1) +
        my_labs() +
        theme(legend.position = "bottom")
```

@@

n_regs <- as.double(nrow(res_sex))
res_sex %>% 
      mutate(
        term = pretty_names(term),
        term = fct_reorder(term, estimate, order_abs(1))
      ) %>% 
      ggplot(aes(y = term, x = estimate, fill = sex, 
            colour = p.value < 0.05/{{n_regs}})) +
        my_vline +
        pval_geom_point() +
        pval_scale() +
        my_fill_scale(n = 2, direction = -1) +
        my_labs()

@@

res_combined %>% 
      mutate(
        score_name = fct_reorder(score_name, estimate, order_abs(2)),
        Regression = ifelse(reg.type == "raw", "Bivariate", "Controlled for age FLB")
      ) %>% 
      ggplot(aes(y = score_name, x = estimate, fill = Regression, 
        color = p.value < 0.05/{{n_regs}})) + 
        geom_vline(xintercept = 0, linetype = "dashed", color = "grey20") +
        pval_geom_point() +
        pval_scale() +
        scale_fill_brewer(type = "qual", palette = 3) +
        my_labs()
```

@@

We can run similar regressions for the parents' generation, using the subsets of respondents who reported their mother's or father's age
and who had no elder siblings. We run *sibling* regressions on these subsets,
controlling for either parent's age at their birth.
Figure \@ref(fig:plot-age-birth-parents) shows
the results. Effect sizes are very similar, whether controlling for
father's or mother's age at respondent's birth or mother's age at respondent's birth. Unlike for the respondents' generation, effect sizes are positively correlated with
the effect sizes from bivariate regressions (father's age at birth: 
$\rho$ `r cor_raw_fath`; mother's age at birth: $\rho$ `r cor_raw_moth`).

```{r plot-age-birth-parents, fig.cap = "Effect sizes on number of siblings controlling for parents' age at birth, eldest siblings"}

@@

res_age_birth_parents %>% 
    filter(! term %in% c("fath_age_birth", "moth_age_birth")) %>% 
    mutate(
      term    = pretty_names(term),
      term    = fct_reorder(term, estimate, order_abs(2)),
      Control = ifelse(control == "fath_age_birth", 
                  "Father's age at birth", 
                  "Mother's age at birth"
                )
    ) %>% 
    ggplot(aes(y = term, x = estimate, fill = Control, 
      colour = p.value < 0.05/{{n_regs}})) + 
      pval_geom_point(alpha = 0.8) +  
      my_vline +
      my_fill_scale(2) +
      pval_scale() +
      my_labs()

```{r calc-age-flb-dv}

drake::loadd(res_age_flb_dv)
drake::loadd(res_age_birth_parents_dv)

compare_flb_dv <- left_join(
        res_age_flb_dv           %>% select(term, estimate),
        res_age_birth_parents_dv %>% select(term, estimate, dep.var),
        by = "term"
      )

cor_flb_fath <- with(compare_flb_dv %>% filter(dep.var == "fath_age_birth"),
        cor(estimate.x, estimate.y)
      )
cor_flb_moth <- with(compare_flb_dv %>% filter(dep.var == "moth_age_birth"),
        cor(estimate.x, estimate.y)
      )
```

These results suggest that polygenic scores may directly correlate with
age at first live birth. Figure \@ref(fig:plot-age-flb-dv) plots
estimated effect sizes from bivariate regressions for respondents, and
Figure \@ref(fig:plot-age-birth-parents-dv) does the same for their parents. 
Effect sizes are reasonably large. They are also very highly correlated across 
generations. Effect sizes of PGS on father's age at own birth, and on own age
at first live birth, have a correlation of `r cor_flb_fath`; for mother's age
and own age it is `r cor_flb_moth`

```{r plot-age-flb-dv, fig.cap = "Effects of polygenic scores on age at first live birth"}

n_regs <- as.double(nrow(res_age_flb_dv))

res_age_flb_dv %>% 
      mutate(
        term = pretty_names(term),
        term = fct_reorder(term, estimate, order_abs(1))
      ) %>% 
      ggplot(aes(estimate, term, colour = p.value < 0.05 / {{n_regs}})) + 
      my_vline +
      pval_geom_point(fill = "steelblue4") +
      pval_scale() + 
      my_labs()

```{r plot-age-birth-parents-dv, fig.cap = "Effects of polygenic scores on parents' age at own birth, eldest siblings"}

n_regs <- as.double(nrow(res_age_birth_parents_dv))

res_age_birth_parents_dv %>% 
      mutate(
        term = pretty_names(term),
        term = fct_reorder(term, estimate, order_abs(1)),
        "Dependent variable" = dplyr::recode(dep.var, 
          "fath_age_birth" = "Father's age", 
          "moth_age_birth" = "Mother's age"
        )
      ) %>% 
      ggplot(aes(estimate, term, fill = `Dependent variable`, 
        colour = p.value < 0.05 / {{n_regs}})) + 
        my_vline +
        pval_geom_point() +
        pval_scale() +
        my_fill_scale(n = 2, direction = -1) +
        my_labs()

@@

n_regs <- as.double(nrow(res_partners))
res_partners %>% 
      mutate(
        score_name = pretty_names(score_name),
        score_name = fct_reorder(score_name, estimate, order_abs(1))
      ) %>% 
      ggplot(aes(y = score_name, x = estimate, fill = `N partners`, 
        colour = p.value < 0.05/{{n_regs}})) + 
        my_vline +
        pval_geom_point() +
        facet_wrap(vars(sex)) +
        my_fill_scale(n = 2) +
        pval_scale() +
        theme(legend.position = "bottom") +
        my_labs(x = "", y = "")

@@

drake::loadd(res_edu)
n_regs <- as.double(nrow(res_edu))
res_edu %>% 
      mutate(
        term = pretty_names(term), 
        term = fct_reorder(term, estimate, order_abs(1)),
        "Age left FTE" = fct_relevel(age_fte_cat, "< 16", "16-18", "> 18")
      ) %>% 
      ggplot(aes(y = term, x = estimate, colour = p.value < 0.05/{{n_regs}}, 
        fill = `Age left FTE`)) +
        my_vline +
        pval_geom_point() + 
        pval_scale() +
        my_fill_scale(n = 3, direction = -1) +
        my_labs()

@@

n_regs <- as.double(nrow(res_income))
res_income %>% 
      mutate(
        term = pretty_names(term), 
        term = fct_reorder(term, estimate, order_abs(1)),
        Income = factor(income_cat, 
          labels = c("< £18K", "£18-30K", "£31-51K", "£52-100K", "> £100K")
        )
      ) %>% 
      ggplot(aes(y = term, x = estimate, fill = Income, 
        colour = p.value < 0.05/{{n_regs}})) +
        my_vline +
        pval_geom_point(alpha = 0.8) + 
        my_fill_scale(n = 5, direction = -1) +
        pval_scale() +
        my_labs()
```
